Insert the article title here

### 2021-07-02 fstats_pca_lewontin.tex
merge

\newcommand{\MX}{\mathbf{X}} %uncentered data
\newcommand{\MC}{\mathbf{C}} %centering
\newcommand{\MY}{\mathbf{Y}} %centered data
\newcommand{\MF}{\mathbf{F}_2} %F2-distance matrix
\newcommand{\MFT}{\mathbf{F}_3} %F3-distance matrix
\newcommand{\MP}{\mathbf{P}} % PCs
\newcommand{\ML}{\mathbf{L}} % loadings
\newcommand{\MK}{\mathbf{K}} % Kernel
\newcommand{\MSINGULAR}{\mathbf{\Sigma}} % Singular values matrix
\newcommand{\MEIGEN}{\mathbf{\Lambda}} % Eigenvalue matrix

\newcommand{\MEAN}{\boldsymbol{\mu}} % Kernel
\begin{document}
\section{Introduction}
About 15\% of genetic variation in humans can be explained by population structure \cite{lewontin1972, barbujani1997, rosenberg2003}, but the information contained in these 15\% is sufficient to study the genetic diversity and history in great detail \cite{edwards2003}. For some data sets it is possible to predict an individuals origin at a resolution of a few hundred kilometers \cite{novembre2008, leslie2015}, and direct-to-consumer-genetics companies are using this variation to analyze the genetic data of millions of customers.

@@

>>>>

\newcommand{\MX}{\mathbf{X}} %uncentered data
\newcommand{\MC}{\mathbf{C}} %centering
\newcommand{\MY}{\mathbf{Y}} %centered data
\newcommand{\MF}{\mathbf{F}_2} %F2-distance matrix
\newcommand{\MFT}{\mathbf{F}_3} %F3-distance matrix
\newcommand{\MP}{\mathbf{P}} % PCs
\newcommand{\ML}{\mathbf{L}} % loadings
\newcommand{\MK}{\mathbf{K}} % Kernel
\newcommand{\MSINGULAR}{\mathbf{\Sigma}} % Singular values matrix
\newcommand{\MEIGEN}{\mathbf{\Lambda}} % Eigenvalue matrix
\newcommand{\MEAN}{\boldsymbol{\mu}} % Kernel

\begin{document}
\section{Introduction}
About 15\% of genetic variation in humans can be explained by population structure \cite{lewontin1972, barbujani1997, rosenberg2003, mallick2016}, but the information contained in these 15\% is sufficient to study our genetic diversity and history in great detail \cite{cavalli-sforza1994, stoneking2016, reich2018}. Direct-to-consumer-genetics companies are using this variation to model the ancestry of millions of customers, and for some data sets, it is possible to predict an individuals' place of origin at a resolution of a few hundred kilometers \cite{novembre2008, leslie2015}.

Human genetic structure is also not particularly discrete.

@@

\section{Results}
\subsection{$F_4$}
Using $F_3$-statistics, I showed that we can think of the admixture test as a test of whether the admixed population lies in a particular $n$-ball, and the outgroup $F_3$-statistic can be thought of as a projection of the test populations on the line connecting the outgroup to the reference sample. In this section, I will develop similar interpretation of $F_4$ on PCA-plots, and to investigate sparsity.

First, we investigate the sparsity in the world overview data set: We find that the vast amount of contribution to the statistics comes from the first two PCs (Figure \ref{fig:f4}A). For example, the correlation between $F_4(X, Y, \text{Mozabite}, \text{Yoruba})$ and its approximation using the first two PCs is 99.2\%. To visualize the interpretation of $F_4$ as an angle, we use statistics of the form $F_4(X, \text{Sardinian}; \text{Mozabite}, \text{Yoruba})$, which can be interpreted as the angle between the vectors Mozabite-Yoruba and $X$-Sardinian. In Figure \ref{fig:f4}B, I show the angle based on two (blue), ten (green) and all PCs $red$. I find that for most Asian and American populations the angle is very close to 90\textdegree, as would be expected if the variation between African and non-African populations is mostly orthogonal. On the other hand, if $X$ is an African population, the angle is lower, and much less well approximated. This demonstrates that this PCA-plot likely does not model within-African population structure adequately.
